---
title: "Webinar-4 Nov. 17, 2024"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Execute the following cells to load the libraries
```{r}
library(ggplot2)
library(dplyr)
```

#### Set ggplot theme for plotting
```{r}
My_Theme = theme(axis.text.x = element_text(size = 9),
   axis.text.y = element_text(size = 9),
   axis.title.x = element_text(size = 11),
   axis.title.y = element_text(size = 11),
   plot.title = element_text(size = 12, hjust = 0.5, face = "bold"))
```

#### Load the food texture dataset
```{r}
#file = 'http://openmv.net/file/food-texture.csv'
file = 'Data/food-texture.csv'
foodData = read.csv(file, header = TRUE, row.names = 1, stringsAsFactors = FALSE)
head(foodData, 2)
str(foodData)
```

#### Modify Crispy column to reflect high (0) or low (1) crispiness
```{r}
foodData = foodData %>% mutate(Crispylevel = ifelse(Crispy > 11, 'High', 'Low'))
```

##### Change Crispy and Crispylevel columns to factor (categorical) type
```{r}
# Continuous features -> Oil, Density, Hardness, Fracture
# Categorical features -> Crispy ('9 levels, 7 through 15') and Crispylevel (2 levels High and Low)
# Crispy is a categorical feature with an order
str(foodData)
categorical_cols = c('Crispy', 'Crispylevel')
foodData[categorical_cols] = lapply(foodData[categorical_cols], as.factor)
str(foodData)
```

#### Visualize the Oil Percentage feature (continuous) using a histogram
```{r}
p = ggplot(data = foodData) + 
  geom_histogram(aes(x = Oil), binwidth = 1, color = 'steelblue') +
  labs(x = 'Oil Percentage',
       y = 'Frequency',
       title = 'Distribution of Oil Percentage') +
  My_Theme
p
```

#### Visualize the Oil Percentage feature (continuous) for Crispy value 10 food items using a histogram
```{r}
p = ggplot(data = foodData %>% filter(Crispy == 10)) + 
  geom_histogram(aes(x = Oil), binwidth = 1, color = 'steelblue') +
  labs(x = 'Oil Percentage',
       y = 'Frequency',
       title = 'Distribution of Oil Percentage for Cripsy 10') +
  My_Theme
p
```

#### Visualize the Density feature (continuous) using a histogram
```{r}
p = ggplot(data = foodData) + 
  geom_histogram(aes(x = Density), binwidth = 100, color = 'steelblue') +
  labs(x = 'Density',
       y = 'Frequency',
       title = 'Distribution of Density') +
  My_Theme
p
```

#### Scatter plot between Oil Percentage and Density (relating two continuous variables)
```{r}
p = ggplot(data = foodData) +
  geom_point(aes(x = Density, y = Oil), color = 'blue') +
  labs(x = 'Density (kg/m^3)',
       y = 'Oil Percentage',
       title = 'Oil Percentage vs. Density') +
  My_Theme
p
```

#### Scatter plot between OilPercentage and Density color coded with Crispylevel
```{r}
p = ggplot(data = foodData) +
  geom_point(aes(x = Density, y = Oil, color = Crispylevel)) +
  labs(x = 'Density (kg/m^3)',
       y = 'Oil Percentage',
       title = 'Oil Percentage vs. Density') +
  My_Theme
p
```

#### Smooth line plot using ggplot
```{r}
p = ggplot(data = foodData) +
  geom_smooth(aes(x = Density, y = Oil), se = FALSE) +
  labs(x = 'Density (kg/m^3)',
       y = 'Oil Percentage',
       title = 'Oil Percentage vs. Density') +
  My_Theme
p
```

#### Plot the distribution of Crispy values (categorical variable)
```{r}
p = ggplot(data = foodData) +
  geom_bar(aes(x = Crispy)) +
  labs(x = 'Crispy',
       y = 'Frequency',
       title = 'Distribution of Crispy Values') +
  My_Theme
p
```

#### Component plot of Density values
```{r}
p = ggplot(data = foodData) +
  geom_point(aes(x = 1:nrow(foodData) , y = Density), color = 'blue') +
  geom_hline(yintercept = mean(foodData$Density), color = 'red') +
  labs(x = 'Sample #',
       y = 'Density (kg/m^3)',
       title = 'Component Plot of Density') +
  My_Theme
p
```


#### Summarizing sample data: mean, variance, standard deviation, median, percentiles, and quantiles
```{r}
# Sample mean of Density values
mean(foodData$Density)

# Mean-centered Density values or the deviations
foodData$Density - mean(foodData$Density)

# Squared deviations
(foodData$Density-mean(foodData$Density))^2

# Average squared deviation (Variance of Density)
mean((foodData$Density-mean(foodData$Density))^2)

# Standard deviation of Density
sqrt(mean((foodData$Density-mean(foodData$Density))^2))

# Standardized Density values (Z-scores of Density)
(foodData$Density - mean(foodData$Density)) / sqrt(mean((foodData$Density-mean(foodData$Density))^2))

# Median of Density 
# Mean is pulled towards the direction of the skew and is therefore not a good representative value when the feature has a skewed distribution
median(foodData$Density)

# The 0.5th quantile is also the 50th percentile which is also referred to as the media
quantile(foodData$Density, 0.5) # get the Density value such that to the left of it we have 50% of the Density values

# The 0.25th quantile of the 25th percentile
quantile(foodData$Density, 0.25) # get the Density value such that to the left of it we have 25% of the Density values

# The 0.75th quantile of the 75th percentile
quantile(foodData$Density, 0.75) # get the Density value such that to the left of it we have 75% of the Density values

# Interquartile range (IQR) = 25th percentile to 75th percentile
c(quantile(foodData$Density, 0.25), quantile(foodData$Density, 0.75))
```

#### Visualize the Density feature using a box plot
```{r}
p = ggplot(data = foodData) +
  geom_boxplot(aes(y = Density)) +
  labs(y = 'Density',
       Title = 'Distribution of Density') +
  My_Theme 
  #coord_flip()
p
```

#### Visualize the Oil percentage feature for different Crispylevel values using box plots
```{r}
p = ggplot(data = foodData) +
  geom_boxplot(aes(x = Crispylevel, y = Density)) +
  labs(y = 'Density',
       Title = 'Distribution of Density') +
  My_Theme 
  #coord_flip()
p
```

#### Load the bodyfat dataset
```{r}
# Load the bodyfat dataset
file = 'Data/bodyfat.csv'
df = read.csv(file, header = TRUE, row.names = 1, stringsAsFactors = F)
str(df)
head(df, n = 2)
```

#### Create lists of categorical and continuous features
```{r}
# Create lists of categorical and continuous features
features = colnames(df)
categorical_features = c(?)
continuous_features = features[?]
```

#### Convert categorical columns to correct type
```{r}
# Convert categorical columns to correct type
df[categorical_features] = lapply(df[categorical_features], as.factor)
str(df)
```

#### Plot the distribution of males and females in the dataset
```{r}
# Plot the distribution of males and females in the dataset
p = ggplot(data = ?) +
  geom_bar(aes(x = ?), color = ?, fill = ?) +
  labs(x = 'Gender',
       y = 'Frequency',
       title = 'Distribution of Genders')
p
#p+coord_flip()
```

#### Plot a component plot of weight values
```{r}
# Plot a component plot of weight values
p = ggplot(data = ?) +
  geom_point(aes(x = ?, y = ?), color = 'blue') +
  labs(x = 'Patient #',
       y = 'Weight (kg)',
       title = 'Component Plot of Weights')
p
```

#### Plot a histogram of weight values
```{r}
# Plot a histogram of weight values
mean_weight = mean(df$Weight)
sd_weight = sd(df$Weight)
p = ggplot(data = df) +
  geom_histogram(aes(x = ?, y = ?),
  breaks = seq(mean_weight-3*sd_weight, mean_weight+3*sd_weight, by = 2.0),
  colour = 'black', fill = 'steelblue', alpha = 0.4) +
  labs(x = 'Weight (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Weights')
p
```

#### Plot a histogram of height values
```{r}
```

#### Plot a histogram of bodyfat values
```{r}
```

#### Plot a histogram of weights for both genders
```{r}
# Plot a histogram of weights for both genders
p = ggplot(data = ?) +
  geom_histogram(aes(x = ?, y = ?, color = ?), alpha = ?) +
  labs(x = 'Weight (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Weights for Males and Females')
p
```

#### Plot a boxplot of bodyfat values
```{r}
# Plot a boxplot of bodyfat values
p = ggplot(data = ?) +
  geom_boxplot(aes(y = ?)) +
  labs(title = 'Bodyfat Distribution')
p
```

#### Plot a boxplot of bodyfat values for both genders
```{r}
# Plot a boxplot of bodyfat values for both genders
p = ggplot(data = ?) +
  geom_boxplot(aes(x = ?, y = ?)) +
  labs(title = 'Bodyfat Distribution for Males and Females')
p
```

#### Introduce a new categorical column in the dataframe called 'Weightlevel' with the following levels: (1) low (< 25th percentile of weights) (2) medium (in the IQR of weights) (3) high (>75th percentile of weights)
```{r}
# Introduce a new categorical column in the dataframe called weight level with the following levels: (1) low (< 25th percentile of weights) (2) medium (in between the 25th and 75th percentiles of weights) (3) high (> 75th percentile of weights)
df['Weightlevel'] = df['Weight'] %>% mutate
```

#### Plot a boxplot of bodyfat values for both genders across the three Weightlevel groups
```{r}
# Plot a boxplot of bodyfat values for both genders both genders across the three Weightlevel groups
````
















